Space Matters: extending sensitivity analysis to initial spatial conditions in geosimulation models

### 2019-08-13 Response_to_Reviewers.tex

  $\rightarrow$ We agree with this comment and recall that urban growth processes are not the main focus of the paper, they are here an instantiation of a given $\gamma$ generator, which could have been replaced by others such as network models, procedural modeling, etc. The position and discussion of the paper was amended to precise that (see comment A).

  %it's not the point of the paper, only the metaphore of the spatial generator. + supprimer la ref à "urban growth" dans le texte et dire d'inspiration urban science mais dont seules les grilles finales nous intéressent. On aurait pu utiliser des generator issus de graphes, de percolation etc.
  %The core of the paper is not this, it's the approach of including gamma and applying to mu to understand systems better, whatever gamma and whatever mu in whatever knowledge domain.
  %\item in discussion, mention that it's one isolated example of geometry but other cases could be thought of as well, using networks and graphs, but also empirical shape files.

  \medskip

  \item \textit{The approach is reductionist and has its strength in reflecting growth at an abstract level nevertheless. The concluding remark in paragraph 5.1 “With the Schelling experiment, we found … that polycentric and discontinuous cities appear systematically more segregated than compact cities …” is elusive. Either it is a trivial fact, simply because polycentric cities have more centres and thus, by nature, more opportunities from which segregation can originate. Or, the results can be used as a starting point to dig deeper in geographical reasons, taking social, economic, cultural, and political, but also topographical issues into consideration. This geographical reasoning, however, is outside of the scope of this paper. }
  %\comment[FL]{j'ai pas compris pourquoi la phrase du reviewer avait été mise en commentaire, j'ai remis}
 %\comment[JR]{I rather agree that it is quite reductionist, and that the discussion may be valid only at very abstract levels (on which we were not explicit maybe ?) - maybe make these conclusions more moderate} \comment[MLT]{I kind of disagree with the idea that it is straightforward that centralised cities produce more segregation as it is not a direct parameter of the model but an output. Plus: what argument would he or she have to state that a big center would be automatically less segregated that several smaller ones? And I believe that explaining why polycentric cities are more segregated than others is out of the scope of the paper. But we could add his or her comment in the discussion section, saying that we hove we have convinced people to add spatial sensitivity analysis during the exploration of their model and that specialist of the thematics would see the opportunity to dig deeper into our results?} \comment[FK]{du coup je fais une proposition de réponse ci-dessous en reprenant ces différents éléments}

  $\rightarrow$ We understand this statement, however we wish to underline that our objective in the discussion of the Schelling results was not to establish a general result on the links between polycentricity and segregation, which are indeed very complex and multidimensional as advocated by the reviewer. Instead, we aimed at showing what kind of results could be obtained using the framework we suggest. If we do not prove that in general, and in empirical context, polycentricity creates more segregation, we obtained, in the modelling context, under Schelling assumptions, that polycentric geometric structures are more often that not associated with segregated patterns, a result which we do not believe to be solely resulting from the fact that more centers leads to more opportunities of segregation (which is true), because it seems also true that monocentric structures, with higher densities, can more easily accommodate heterogeneity of agents within cells, and hence a more patchy type of segregation which can for instance being observed in London. We included a comment in the discussion section, saying that we hope we have convinced people to add spatial sensitivity analysis during the exploration of their model and that specialist of the thematics would see the opportunity to dig deeper into our results \comment[FL]{à la relecture de l'article s'assurer que cette prise de recul est bien explicitée}

### 2019-08-13 paper.tex
Update on Overleaf.

\subsubsection{Spatial processes}
In 'real-life', different spatial configurations of people in a city tend to be associated with different distributions of income, carbon emissions, educational outcomes, etc. For example, \citet{wheeler2006urban} shows that, in the US, sprawling cities are more unequal than their compact counterparts with respect to income. Dynamically, sprawl in American cities consists in the addition of new developments which have been occupied by different groups of population, resulting in a concentration of the wealthy in suburban pockets and of pockets of poverty in the inner city area \citep{jargowsky2002sprawl}. Similarly, in terms of pollution for example, \citet[p.173]{schwanen2001travel} show that \textit{``deconcentration of urban land uses encourages driving and discourages the use of public transport as well as cycling and walking''}\comment[FL]{je ne comprends pas ce point : est-ce lié à un modèle de simulation?}. These effects of geographical space on social systems correspond to processes to be included in geosimulation models.\\

>>>>

\subsubsection{Spatial processes}
In 'real-life', different spatial configurations of people in a city tend to be associated with different distributions of income, carbon emissions, educational outcomes, etc. For example, \citet{wheeler2006urban} shows that, in the US, sprawling cities are more unequal than their compact counterparts with respect to income. Dynamically, sprawl in American cities consists in the addition of new developments which have been occupied by different groups of population, resulting in a concentration of the wealthy in suburban pockets and of pockets of poverty in the inner city area \citep{jargowsky2002sprawl}. Similarly, in terms of pollution for example, \citet[p.173]{schwanen2001travel} show that \textit{``deconcentration of urban land uses encourages driving and discourages the use of public transport as well as cycling and walking''}. These effects of geographical space on social systems correspond to processes to be included in geosimulation models.\\
